---
title: "Project 2"
output:
  html_document:
    df_print: paged
---

# Load Packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(vip)
library(parallel)
library(doParallel)
library(xgboost)
```

# Load Initial Data

```{r, warning=FALSE, message=FALSE}

raw_data <- read_csv("../Project 2/boston_train.csv") %>%
  clean_names()

head(raw_data)

zip_code <- read_csv("../Project 2/zips.csv")

head(zip_code)

holdout <- read_csv("../Project 2/boston_holdout.csv") %>%
  clean_names()


```

#  First Skim

```{r}
# raw_data %>%
#   skim_without_charts()
```

after skimming,
PID, ZIPCODE should be transformed into categorical

zip - already there
we can just remove PID, ZIPCODE

train - 33 variables
holdout - 32 variables

# Transformation and Merge

```{r}
raw_data$yr_built <- as.character(raw_data$yr_built)
raw_data$yr_remod <- as.character(raw_data$yr_remod)

holdout$yr_built <- as.character(holdout$yr_built)
holdout$yr_remod <- as.character(holdout$yr_remod)

raw_data %>%
  select(-pid, -zipcode) %>%
  mutate_if(is.character, factor) %>%
  clean_names() -> raw_data

raw_data %>%
  skim_without_charts()

holdout %>%
  select(-pid, -zipcode) %>%
  mutate_if(is.character, factor) %>%
  clean_names() -> holdout

holdout %>%
  skim_without_charts()
```
# Summary

## Numeric Summary

```{r}

# Defining the function
my.num.summary <- function(x){
  c(n=NROW(x),
    Distinct=n_distinct(x),
    Mean=mean(x),
    SD=sd(x),
    Median=median(x),
    Min=min(x),
    Max=max(x), 
    "# Missing"=n_missing(x),
    "% Missing"=n_missing(x)/NROW(x)
    )
  }

# identifying numeric columns
num <- sapply(raw_data, is.numeric)

# applying the function to numeric columns only
num <- sapply(raw_data[, num], my.num.summary)

# transpose
num_summary <- as.data.frame(t(num))

htmlTable::htmlTable(num_summary)

```

## Categorical summary

```{r}
my.cat.summary <- function(x){
  c(Datatype=class(x),
    n=NROW(x),
    Distinct=n_distinct(x),
    "Pct Distinct"=n_distinct(x)/NROW(x),
    "# Missing"=n_missing(x),
    "% Missing"=n_missing(x)/NROW(x)
    )
}

cat <- sapply(raw_data, is.factor)

cat <- sapply(raw_data[, cat], my.cat.summary)

# transpose
cat_summary <- as.data.frame(t(cat))

htmlTable::htmlTable(cat_summary)
```

# target summary

```{r}
options(scipen = 999)
raw_data %>%
  group_by(av_total) %>%
  summarise(n=n()) %>%
  ungroup() %>%
  ggplot(aes(av_total)) +
  geom_histogram(bins = 30) +
  labs(title="target distribution")
```



# exploration

```{r}
# OWN_OCC
raw_data %>%
  group_by(own_occ, av_total) %>%
  summarise(n=n()) %>%
  ggplot(aes(av_total,fill=own_occ)) +
  geom_histogram(bins=30) +
  labs(title="Does owner-occupied homes have a higher assessed value")

# YEAR_REMOD
raw_data %>%
  ggplot(aes(yr_remod,av_total)) +
  geom_point() +
  labs(title="Does year remod impact assessed value") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.4))

```

# Initial screening and exploration

```{r}
# living area
raw_data %>%
  ggplot(aes(living_area, av_total)) +
  geom_point() +
  labs(title="bigger living area leads to higher value")

# land_sf
raw_data %>%
  ggplot(aes(land_sf, av_total)) +
  geom_point() +
  labs(title="Parcel’s land area have little impact on value")

# total_rooms
raw_data %>%
  ggplot(aes(r_total_rms, av_total)) +
  geom_point() +
  labs(title="Total rooms count slightly impact value")

# overall condition
raw_data %>%
  group_by(r_ovrall_cnd) %>%
  summarize(avg = mean(av_total),
            median = median(av_total))

# median income
raw_data %>%
  ggplot(aes(median_income, av_total)) +
  geom_point() +
  labs(title="does median income affect value?")
```


# Split and Defline K-fold

here we want to split our data into 70/30 training to test 
THEN apply k-fold to our training data
you'll use the kfold grid search to find optimal parameters, which you will then apply a final fit 


```{r}
# -- set a random seed for repeatablity 
# set.seed(123)
# 
# # -- train test split 
# train_test_spit<- initial_split(raw_data, prop = 0.7)
# 
# train <- training(train_test_spit)
# test  <- testing(train_test_spit)
# 
# # -- k-fold your training data 
# train_cv_folds <- vfold_cv(train, v=10)
# 
# 
# sprintf("Train PCT  : %1.2f%%", nrow(train)/ nrow(raw_data) * 100)
# sprintf("Test  PCT  : %1.2f%%", nrow(test)/ nrow(raw_data) * 100)
# sprintf("Kfold Count: %d", nrow(train_cv_folds))


```



# Define recipe

```{r}

# bos_recipe <- 
#   recipe(av_total ~ ., data=train) %>%
#   # step_rm(pid) %>%
#   step_medianimpute(all_numeric(), -all_outcomes()) %>%
#   step_novel(all_nominal(), -all_outcomes()) %>%
#   step_unknown(all_nominal(), -all_outcomes()) %>%
#   step_dummy(all_nominal(), -all_outcomes())

```

# Define models
for both XGB and RF

```{r}
# xgb_model <- boost_tree(
#   mtry = tune(),
#   trees  = tune(),
#   min_n = tune(),
#   learn_rate = tune()
#   ) %>% 
#     set_engine("xgboost") %>% 
#     set_mode("regression")
# 
# rf_model <- rand_forest(
#   trees = tune(),
#   min_n = tune()
#   ) %>%
#   set_engine("ranger", importance = "permutation") %>%
#   set_mode("regression")
```

# Man learning 
xD

```{r}
# xgb_model <- boost_tree(
#   mtry = 20,
#   trees  = 1000,
#   min_n = 7,
#   learn_rate = 0.05
#   ) %>% 
#     set_engine("xgboost") %>% 
#     set_mode("regression")
# 
# xgb_workflow <- workflow() %>%
#   add_recipe(bos_recipe) %>%
#   add_model(xgb_model)
# 
# xgb_final_wf <- xgb_workflow
#   # finalize_workflow(xgb_best)
# 
# xgb_final_fit <- xgb_final_wf %>%
#   fit(train)
# 
# # score
# predict(xgb_final_fit, train) %>%
#   bind_cols(., train) -> train_scored_xgb
# 
# predict(xgb_final_fit, test, type="numeric") %>%
#   bind_cols(.,test) -> test_scored_xgb
# 
# # Metrics: Train and Test 
# train_scored_xgb %>%
#   mutate(part="train") %>%
#   bind_rows(test_scored_xgb %>%
#               mutate(part = "test")) %>%
#   group_by(part) %>%
#   metrics(av_total, estimate = .pred) %>%
#   filter(.metric %in% c('rmse','rsq')) %>%
#   pivot_wider(names_from = .metric, values_from=.estimate) %>%
#   print()
```


# Define workflows

```{r}
# xgb_workflow <- workflow() %>%
#   add_recipe(bos_recipe) %>%
#   add_model(xgb_model)
# 
# rf_workflow <- workflow() %>%
#   add_recipe(bos_recipe) %>%
#   add_model(rf_model)
```

# set up tuning grids

```{r}
# --  xgb grid
# xgb_tune_grid <- grid_regular(mtry(c(16,20)),
#                               trees(c(1000,1000)),
#                               min_n(c(6,10)),
#                               learn_rate(c(-2,-2)),
#                               levels = 5)
# 
# print(xgb_tune_grid)
# 
# # -- rf grid 
# rf_tune_grid <- grid_regular(trees(c(1000,1000)),
#                              min_n(),
#                              levels = 5)
# 
# print(rf_tune_grid)
```

# Kick the fit off with parallel computation

```{r, warning=FALSE}
# all_cores <- detectCores(logical = TRUE)
# sprintf("# of Logical Cores: %d", all_cores)
# 
# cl <- makeCluster(all_cores)
# 
# registerDoParallel(cl)
# 
# # -- XGB
# xgb_tuning_results <- xgb_workflow %>%
#   tune_grid(
#     resamples = train_cv_folds,
#     grid = xgb_tune_grid
#   )
# 
# # -- rf
# rf_tuning_results <- rf_workflow %>%
#   tune_grid(
#     resamples = train_cv_folds,
#     grid = rf_tune_grid
#   )
```

# collect and prepare for eval

```{r}

# -- xgb

# xgb_tuning_results %>%
#   collect_metrics() %>%
#   mutate_if(is.numeric, round, 3)
# 
# xgb_tuning_results %>%
#   show_best("rmse") %>%
#   print()
# 
# # -- rf
# 
# rf_tuning_results %>%
#   collect_metrics() %>%
#   mutate_if(is.numeric, round, 3)
# 
# rf_tuning_results %>%
#   show_best("rmse") %>%
#   print()
```



# Select the best and refit

```{r, warning=FALSE}

# -- xgb
# xgb_tuning_results %>%
#   show_best("rmse") %>%
#   print()
# 
# xgb_best <- xgb_tuning_results %>%
#   select_best("rmse")
# 
# xgb_final_wf <- xgb_workflow %>%
#   finalize_workflow(xgb_best)
# 
# xgb_final_fit <- xgb_final_wf %>%
#   fit(train)
# 
# # -- rf
# rf_tuning_results %>%
#   show_best("rmse")
# 
# rf_best <- rf_tuning_results %>%
#   select_best("rmse")
# 
# rf_final_wf <- rf_workflow %>%
#   finalize_workflow(rf_best)
# 
# rf_final_fit <- rf_final_wf %>%
#   fit(train)
```

# evaluate

```{r}
# -- xgb
# xgb_final_fit %>%
#   pull_workflow_fit() %>%
#   vip(n = 20)
# 
# # -- rf
# rf_final_fit %>%
#   pull_workflow_fit() %>%
#   vip(n = 20)
```

```{r}
# regression_eval <- function(model){
# 
# # -- score training  
#   predict(model, train) %>%
#     bind_cols(.,train)-> train_scored 
# 
#   predict(model, test, type="numeric") %>%
#     bind_cols(.,test)-> test_scored 
#   
#   # -- Metrics: Train and Test 
#   train_scored %>% 
#     mutate(part="train") %>%
#     bind_rows(test_scored %>% mutate(part = "test")) %>%
#     group_by(part) %>%
#     metrics(av_total, estimate = .pred) %>%
#     filter(.metric %in% c('rmse','rsq')) %>%
#     pivot_wider(names_from = .metric, values_from=.estimate) %>%
#     print()
# 
# }
#   
# regression_eval(xgb_final_fit)
# regression_eval(rf_final_fit)
```
1st run: learn_rate -> 0.100 is good, increase trees by 100 could lower mean by ~400
2nd run: tree_depth close to 4 is good 
3rd run: xgb -> min_n(30,40) mtry(15,20) rf -> min_n=2 mtry=20
4th: mtry(16,20), learn rate(-2,-0.8) -> best 52334
5rd: mtry(c(16,20)), min_n(c(6,10)), learn_rate(c(-2,-2))


# predict holdout

```{r}
# predict(xgb_final_fit, holdout) %>%
#   bind_cols(., holdout %>%
#               select(pid)
#             ) %>%
#   select(pid, av_total=.pred) -> kaggle_submit
# 
# write_csv(kaggle_submit, "proj2_submission4.csv")
```

# Let's see what would happen if I follow every bit of details in the sample code ...

a few things here:
I dont see the yr_remod is categorical - numeric does not make sense
why would tuning 2 parameters yield a far better result than tuning 4 to 5?

```{r}

# load the data, clean the names, make factors 

df <-read_csv("../Project 2/boston_train.csv") %>%
  clean_names() %>%
  mutate_if(is.character,factor)

holdout <-read_csv("../Project 2/boston_holdout.csv") %>%
  clean_names() %>%
  mutate_if(is.character,factor)

# -- set a random seed for repeatablity 
set.seed(42)

# -- train test split 
train_test_spit<- initial_split(df, prop = 0.7)

train <- training(train_test_spit)
test  <- testing(train_test_spit)

# -- k-fold your training data 
train_cv_folds <- vfold_cv(train, v=5)

sprintf("Train PCT  : %1.2f%%", nrow(train)/ nrow(df) * 100)
sprintf("Test  PCT  : %1.2f%%", nrow(test)/ nrow(df) * 100)
sprintf("Kfold Count: %d", nrow(train_cv_folds))


# -- define recipe 
bos_recipe <- 
  recipe(av_total ~ ., data=train) %>%
  step_rm(pid) %>%
  step_medianimpute(all_numeric(), -all_outcomes()) %>%
  step_unknown(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes())

# -- define models 
xgb_model <- boost_tree(
    trees  = tune(),
    learn_rate = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("regression")

rf_model <- rand_forest(
  trees = tune(),
  min_n = tune()) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("regression")

# -- define workflows 
xgb_workflow <- workflow() %>%
  add_recipe(bos_recipe) %>%
  add_model(xgb_model)

rf_workflow <- workflow() %>%
  add_recipe(bos_recipe)  %>%
  add_model(rf_model)

# --  xgb grid
xgb_tune_grid <- grid_regular(trees(c(100,1000)),
                              learn_rate(),
                              levels = 10)

print(xgb_tune_grid)

# -- rf grid 
rf_tune_grid <- grid_regular(trees(c(100,1000)),
                          min_n(),
                          levels = 5)

print(rf_tune_grid)

all_cores <- detectCores(logical = TRUE)
sprintf("# of Logical Cores: %d", all_cores)

cl <- makeCluster(all_cores)

registerDoParallel(cl)

# -- K-Fold XGB  
xgb_tuning_results <- 
  xgb_workflow %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = xgb_tune_grid,
    #control = control_resamples(save_pred = TRUE)
    )

# -- k-fold RF
rf_tuning_results <- 
  rf_workflow %>% 
  tune_grid(
    resamples = train_cv_folds,
    grid = rf_tune_grid,
    #control = control_resamples(save_pred = TRUE)
    )


# -- evaluate your Hyper Parameter K-Folds
xgb_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err)) 
  
xgb_tuning_results %>%
  show_best("rmse") %>%
  print()

xgb_best_rmse <- xgb_tuning_results %>%
  select_best("rmse") 

rf_tuning_results %>% 
  collect_metrics() %>%
  mutate_if(is.numeric, round,3) %>% 
  pivot_wider(names_from = .metric, values_from=c(mean, std_err)) 

rf_tuning_results %>%
  show_best("rmse") %>%
  print()

rf_best_rmse <- rf_tuning_results %>%
  select_best("rmse") 

print(xgb_best_rmse)
print(rf_best_rmse)

# -- finalize XBG workflow 
xgb_final_wf <- 
  xgb_workflow %>% 
  finalize_workflow(xgb_best_rmse)

xgb_final_fit  <- 
  xgb_final_wf %>%
  fit(data = train) 

# -- finalize RF workflow 
rf_final_wf <- 
  rf_workflow %>% 
  finalize_workflow(rf_best_rmse)

rf_final_fit  <- 
  rf_final_wf %>%
  fit(data = train) 

xgb_final_fit %>% 
  pull_workflow_fit() %>% 
  vip(40) + labs(title = "XGB var importance")

rf_final_fit %>% 
  pull_workflow_fit() %>% 
  vip(40)+ labs(title = "RF var importance")

regression_eval <- function(model){

# -- score training  
  predict(model, train) %>%
    bind_cols(.,train)-> train_scored 

  predict(model, test, type="numeric") %>%
    bind_cols(.,test)-> test_scored 
  
  # -- Metrics: Train and Test 
  train_scored %>% 
    mutate(part="train") %>%
      bind_rows(test_scored %>% mutate(part = "test")) %>%
    group_by(part) %>%
    metrics(av_total, estimate = .pred) %>%
    filter(.metric %in% c('rmse','rsq')) %>%
    pivot_wider(names_from = .metric, values_from=.estimate) %>%
    print()

}
  
regression_eval(xgb_final_fit)
regression_eval(rf_final_fit)


predict(xgb_final_fit, holdout) %>%
  bind_cols(.,holdout %>% select(pid)) %>%
  select(pid, av_total = .pred) -> sample_submission

write_csv(sample_submission, "sample_submission2.csv")
```

sample submission gives us 51728 on Kaggle - super confusing...

xgb_tune_grid <- grid_regular(trees(c(100,1000)),
                          learn_rate(),
                          levels = 5) -> 50272
                          
xgb_tune_grid <- grid_regular(trees(c(100,1000)),
                          min_n(),
                          learn_rate(),
                          levels = 5) -> 49799 BAD KAGGLE

xgb_tune_grid <- grid_regular(trees(c(100,1000)),
                              tree_depth(),
                              learn_rate(),
                              levels = 10) -> 49727 BAD KAGGLE

# predictions file

for regression problems we want to compare histograms of predicted vs actual av_total 
they should have same / similar shape.

```{r}
options(scipen = 999)
df %>%
  ggplot(aes(av_total)) +
  geom_histogram(bins = 50) +
  labs(title="actual total value histogram")

sample_submission %>%
  ggplot(aes(av_total)) +
  geom_histogram(bins = 50) +
  labs(title="predicted value histogram")
```






